Tree height and hydraulic traits shape growth responses across droughts in a temperate broadleaf forest

### 2020-02-04 manuscript/climate_sensitivity_draft.Rmd
worked on discussion

The development of tree-ring chronologies for the twelve most dominant tree species at our site [@helcoski_growing_2019; @bourg_initial_2013] gave us the sample size to compare historical drought responses across species and associated traits at a single site [see also @elliott_forest_2015]. Concerted measurement of leaf hydraulic traits of emerging importance [@scoffoni_leaf_2014; @bartlett_correlations_2016; @medeiros_extensive_2019] allowed novel insights into the role of hydraulic traits in shaping drought response. The finding that $PLA_{dry}$ and $\pi_{tlp}$ can be useful for predicting drought responses-with both appearing in the top overall model and two of the three top models for each drought (Table 5)-is consistent with studies demonstrating that these are physiologically meaningful traits linked to species distribution along moisture gradients [@medeiros_extensive_2019; @simeone_coupled_2019; @marechaux_drought_2015] (*Rosas et al. 2019, DOI: 10.1111/nph.15684*, [Fletcher et al. 2018](https://bsapubs.onlinelibrary.wiley.com/doi/full/10.1002/ajb2.1164)). Importantly, this finding indicates that both $PLA_{dry}$ and $\pi_{tlp}$, which can be measured relatively easily  [@bartlett_rapid_2012; @scoffoni_leaf_2014], hold promise for predicting drought growth responses across species. The importance of linking species' traits to drought responses increases with tree species diversity; whereas it is feasible to study drought responses for all dominant species in most boreal and temperate forests (e.g., this study), this becomes difficult to impossible for species that don't form annual rings, and for diverse tropical forests. Although progress is being made for the tropics [@schongart_dendroecological_2017], a full linkage hydraulic traits to drought responses would be invaluable for forecasting how little-known species and whole forests will respond to future droughts (**REFS?--Nate Swenson papers? there was a recent review in TREE**).

>>>>

The development of tree-ring chronologies for the twelve most dominant tree species at our site [@helcoski_growing_2019; @bourg_initial_2013] gave us the sample size to compare historical drought responses across species and associated traits at a single site [see also @elliott_forest_2015]. Concerted measurement of leaf hydraulic traits of emerging importance [@scoffoni_leaf_2014; @bartlett_correlations_2016; @medeiros_extensive_2019] allowed novel insights into the role of hydraulic traits in shaping drought response. The finding that $PLA_{dry}$ and $\pi_{tlp}$ can be useful for predicting drought responses of tree growth (Tables 1,4,5) is both novel and consistent with previous studies linking these traits to habitat and drought tolerance. Previous studies have demostrated that $\pi_{tlp}$ and $PLA_{dry}$ are physiologically meaningful traits linked to species distribution along moisture gradients [@medeiros_extensive_2019; @simeone_coupled_2019; @marechaux_drought_2015] (*Rosas et al. 2019, DOI: 10.1111/nph.15684*, [Fletcher et al. 2018](https://bsapubs.onlinelibrary.wiley.com/doi/full/10.1002/ajb2.1164)), and our findings indicate that these traits also influence drought responses. Furthermore, the observed linkage of $\pi_{tlp}$ to $Rt$ in this forest aligns with observations in the Amazon that $\pi_{tlp}$ is higher in drought-intolerant that drought-tolerant plant functional types and adds support to the idea that this trait is useful for categorizing and representing species' drought responses in models (*Powell et al. 2016, DOI: 10.1111/gcb.13731*). Because both $PLA_{dry}$ and $\pi_{tlp}$, which can be measured relatively easily  [@bartlett_rapid_2012; @scoffoni_leaf_2014], they hold promise for predicting drought growth responses across species. The importance of linking species' traits to drought responses increases with tree species diversity; whereas it is feasible to study drought responses for all dominant species in most boreal and temperate forests (e.g., this study), this becomes difficult to impossible for species that don't form annual rings, and for diverse tropical forests. Although progress is being made for the tropics [@schongart_dendroecological_2017], a full linkage hydraulic traits to drought responses would be invaluable for forecasting how little-known species and whole forests will respond to future droughts (*Powell et al. 2016, DOI: 10.1111/gcb.13731*).
